Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-10-04 paper/sections/metis-language-and-proofs.tex
[ ref.bib ] added a new one ref.

In this paper, we ported a subset of the \Metis' inference rules to
\Agda, the propositional fragmented without equality.
That allows us to justify step-by-step the proofs delivered in
\TSTP format generated by the prover \Metis.

@@

This lack of the \Metis' documentation is an important matter for
us in the proof-reconstruction and certainty for other implementors
(see a comment related on the presenting of \name{Sledgehammer}
in \cite{paulson2007source} and \cite{Farber2015}). We faced
the ambiguity and lack of clearness of some rules like \canonicalize
or \simplify.
For example, we believe by applying these last
inferences, \Metis summarized many essential steps to deduce a
formula from a list of formulas.
\improvement{Add an example here to show a deduction where \Metis do
a lot between two steps.}
That inference's behavior is a drawback since, our approach
to reconstruct proofs needs take into account any assumption or
consideration as ordering for literals, application of
simplification theorems, reductions, among other valid
transformations.

\unsure{Should I give an Example?}

In this part, our main purpose consists of giving a response to
these rules looking at the \Metis' source code while inspecting a
representative set of solutions of \CPL problems
\cite{Prieto-Cubides2017} generated by \Metis.
This is an attempt to fulfill the \Metis's documentation, by
describing each rule from our perspective. Later on, in Section
\ref{sec:proof-reconstruction}, we present a formal description of
such rules.

@@

\textit{Clausification.} The \clausify rule transforms a
propositional formula in clausal normal form or \CNF, a conjunction
of clauses. Where a clause is the disjunction of zero or more
literals and a literal is an atom (positive literal) or a negation
of an atom (negative literal). This conversion is not unique, and
\Metis performs this transformation using its criteria.
We refer the reader to see Theorem~\ref{thm:thm-clausify}.

\improvement{We should mention that it was indeed a problem for the
proof-reconstruction? and refer the section next?}

@@

\textit{Normalization.} \Metis uses the \canonicalize rule to
normalize a propositional formula that comes from an axiom or a
definition. It often converts each expression in one of its normal
form: the conjunctive normal form, the negative normal form or the
disjunctive normal form. Next, this rule simplifies the formula
with some definitions, applying recursively on the structure's
formula conjunctions and disjunctions simplifications to remove
tautologies (see Fig.~\ref{fig:conjunctive-disjunctive-simpl} for a
list of these theorems. We assume the commutative property for the
conjunction and disjunction connectives).
We refer the reader to see Section~\ref{sssec:normalization} and the
lemmas and theorems described there.

>>>>

In this paper, we ported a subset of the \Metis' inference rules to
\Agda, the propositional fragmented without equality, that allows us to justify step-by-step the proofs delivered in
\TSTP format generated by the prover \Metis.

@@

This lack of the \Metis' documentation represents one of the first barriers to build proof-reconstruction tools since some inference rules
like \canonicalize or \simplify summarize many essential steps to deduce the conclusion, making them hard to justify and emulate
(see a comment related on the presenting of \name{Sledgehammer}
in \citeauthor{paulson2007source}~\cite{paulson2007source} and \citeauthor{Farber2015}~\cite{Farber2015}).

Henceforth, one of our main purposes with this document consists of
giving a response to these rules looking at the \Metis' source code but also reviewing a representative set of solutions generated by
\Metis for the \CPL collection of \TPTP problems~\cite{Prieto-Cubides2017} including some problem from the \Metis repository itself.

In the following, we briefly describe each \Metis' rule from our perspective, but we present a formal description for each rule in Section \ref{sec:proof-reconstruction}.

@@

\textit{Clausification.} The \clausify rule transforms a
propositional formula in clausal normal form, a conjunction
of clauses. Where a \emph{clause} is the disjunction of zero or more
literals and a \emph{literal} is an atom (positive literal) or a
negation of an atom (negative literal). Be aware this kind of conversion between one formula to its clausal normal form is not unique, and \Metis has a customized approach to perform that transformation. We refer the reader to see Theorem~\ref{thm:thm-clausify}.

@@

\textit{Normalization.}
Each application of the clausify rule described above is preceded by
\canonicalize rule. Both rules aim to perform a common task for
automatic theorem provers, clausification, mainly described on paper
by \citeauthor{Sutcliffe1996}~\cite{Sutcliffe1996}.

\Metis uses the \canonicalize rule to
normalize a propositional formula that comes from an axiom or a
definition. It often converts each expression in one of its normal
form: the conjunctive normal form, the negative normal form or the
disjunctive normal form. Next, this rule simplifies the formula
with some definitions, applying recursively on the structure's
formula conjunctions and disjunctions simplifications to remove
tautologies (see Fig.~\ref{fig:conjunctive-disjunctive-simpl} for a
list of these theorems. We assume the commutative property for the
conjunction and disjunction connectives).
We refer the reader to see Section~\ref{sssec:normalization} for the
lemmas and theorems described there.
